### 2021-01-30 WOOT/BadUSBC.tex
Update abs

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\makeatletter
\def\lst@makecaption{%
  \def\@captype{table}%
  \@makecaption
}
\makeatother

\input{./Files/util.tex}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{\tool: Revisiting BadUSB with Type-C}

\author{Anonymous Authors}
%\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%}

\maketitle
\thispagestyle{plain}
\pagestyle{plain}

\begin{abstract}
The security of USB protocol has been paid extensive attention to, despite its wide usage.
Due to the \textit{trust-by-default} characteristics, USB security has caused serious problems.
For example, a well-known firmware attack, BadUSB, performs malicious operations on the victim hosts through disguising normal USB devices as human interface devices like keyboards and mice.
However, BadUSB suffers from several limitations.
Attackers cannot obtain the status of user interface to conduct precise attacks, and they may be detected by multiple effective countermeasures.
In this work, we extended BadUSB to support the new features of USB Type-C and proposed a multi-mode attack model, \tool.
It can obtain UI status to make attacks more precise and effective, and also bypass existing defense methods for BadUSB.
To our best knowledge, \tool is the first attack model utilizing USB Type-C.
To validate the usability and effectiveness, we conducted extensive experiments and a user study on a real-life attack scenario.
10 volunteers are involved in, and 4,796 of 94,058 records obtained by \tool in the user study are related to privacy information.
\fengwei{What results?}
%\fengwei{what countermeasures do we propose? including isolated UI rendering.}
We also proposed reasonable countermeasures for our attack model, including isolated UI rendering, external hardware authorization and distrust-by-default, which may be inspiring for future research on defense methods.
\end{abstract}

\begin{IEEEkeywords}
USB;BadUSB;
\end{IEEEkeywords}

\input{./Files/introduction.tex}
\input{./Files/background.tex}
\input{./Files/related_work.tex}
\input{./Files/badusb.tex}
\input{./Files/experiment.tex}
\input{./Files/countermeasure.tex}
\input{./Files/discussion.tex}
\input{./Files/conclusion.tex}
%\input{./Files/acknowledgement.tex}

\bibliographystyle{IEEEtran}
\bibliography{BadUSBC}
\balance

\end{document}