### 2021-01-30 WOOT/BadUSBC.tex
pass on abstract

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.

\makeatletter
\def\lst@makecaption{%
  \def\@captype{table}%
  \@makecaption
}
\makeatother

\input{./Files/util.tex}

\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}

\begin{document}

\title{\tool: Revisiting BadUSB with Type-C}

\author{Anonymous Authors}
%\author{\IEEEauthorblockN{1\textsuperscript{st} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%\and
%\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
%\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
%\textit{name of organization (of Aff.)}\\
%City, Country \\
%email address or ORCID}
%}

\maketitle
\thispagestyle{plain}
\pagestyle{plain}

\begin{abstract}

The security of USB protocol has been paid extensive attention to, despite its
    wide usage.  Due to the \textit{trust-by-default} characteristics, USB
    security has caused serious problems.  For example, a well-known firmware
    attack, BadUSB, performs malicious operations on the victim hosts through
    disguising normal USB devices as human interface devices like keyboards and
    mice.  However, BadUSB suffers from several limitations.  Attackers cannot
    obtain the status of user interface to conduct precise attacks, and they
    can detected by multiple existing countermeasures.  In this work, we
    extended BadUSB to support the new features of USB Type-C and proposed a
    multi-mode attack model, \tool.  It obtains UI status to make attacks more
    precise and effective, and also bypasses existing defense methods for
    BadUSB.  To the best of our knowledge, \tool is the first attack model
    utilizing USB Type-C.  To validate the usability and effectiveness, we
    conducted extensive experiments and a user study on a real-life attack
    scenario.  10 volunteers are involved in, and 4,796 of 94,058 records
    obtained by \tool in the user study are related to sensitive information.
%\fengwei{What results?} \fengwei{what countermeasures do we propose? including
    %isolated UI rendering.}
We also discussed our recommended countermeasures for our attack model,
    including isolated UI rendering, which may be inspiring for future research
    on defense methods.

\end{abstract}

\begin{IEEEkeywords}
USB;BadUSB;
\end{IEEEkeywords}

\input{./Files/introduction.tex}
\input{./Files/background.tex}
\input{./Files/related_work.tex}
\input{./Files/badusb.tex}
\input{./Files/experiment.tex}
\input{./Files/countermeasure.tex}
\input{./Files/discussion.tex}
\input{./Files/conclusion.tex}
%\input{./Files/acknowledgement.tex}

\balance

\bibliographystyle{IEEEtran}
\bibliography{BadUSBC}

\end{document}